---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(knitr.kable.NA = "")
```

# SMPDS: Modern Pollen Data for climate reconstructionS

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/smpds)](https://CRAN.R-project.org/package=smpds)
[![Codecov test coverage](https://codecov.io/gh/special-uor/smpds/branch/main/graph/badge.svg)](https://codecov.io/gh/special-uor/smpds?branch=main)
[![R-CMD-check](https://github.com/special-uor/smpds/workflows/R-CMD-check/badge.svg)](https://github.com/special-uor/smpds/actions)
<!-- badges: end -->

The goal of smpds is to ...

## Installation

You can install the released version of SMPDS from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("smpds")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("special-uor/smpds")
```
## Example

<!-- This is a basic example which shows you how to solve a common problem: -->

```{r example, eval = TRUE}
library(smpds)
## basic example code
smpds::SMPDSv2 %>%
  dplyr::relocate(DOI, .after = publication) %>%
  dplyr::slice(1:20) %>%
  smpds::rm_na_taxa(cols = 1:120) %>% # Filter taxa without counts
  dplyr::select(1:70) %>% # Select only the metadata (16) + the 50 first taxa
  knitr::kable()
```

#### Plots 
##### Potential Natural Vegetation (PNV)
```{r example-plot, eval =  TRUE, fig.width = 14, fig.height = 8, dpi = 300}
smpds::SMPDSv2 %>%
  smpds::plot_biome()
```

##### Growing Degree Days aboce 0 deg Celsius (`gdd0`)
```{r example-plot-gdd0, eval =  TRUE, fig.width = 14, fig.height = 8, dpi = 300}
smpds::SMPDSv2 %>%
  smpds::plot_gdd()
```

##### Mean Temperature of the COldest month (`mtco`)
```{r example-plot-mtco, eval =  TRUE, fig.width = 14, fig.height = 8, dpi = 300}
smpds::SMPDSv2 %>%
  smpds::plot_mtco()
```

##### Moisture Index (`mi`)
```{r example-plot-mi, eval =  TRUE, fig.width = 14, fig.height = 8, dpi = 300}
smpds::SMPDSv2 %>%
  smpds::plot_mi()
```

## Code of Conduct
  
  Please note that the SMPDS project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
