---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# smpds

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/smpds)](https://CRAN.R-project.org/package=smpds)
[![Codecov test coverage](https://codecov.io/gh/special-uor/smpds/branch/main/graph/badge.svg)](https://codecov.io/gh/special-uor/smpds?branch=main)
[![R-CMD-check](https://github.com/special-uor/smpds/workflows/R-CMD-check/badge.svg)](https://github.com/special-uor/smpds/actions)
<!-- badges: end -->

The goal of smpds is to ...

## Installation

You can install the released version of smpds from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("smpds")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("special-uor/smpds")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example, eval = FALSE}
library(smpds)
## basic example code
```

## Code of Conduct
  
  Please note that the smpds project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
