---
title: "A1. CRU TS 4.04"
output: 
  github_document:
    pandoc_args: --webtex
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = FALSE,
                      comment = "#>",
                      fig.path = "man/figures/cru-ts-4.04-",
                      out.width = "100%",
                      dpi = 300, 
                      fig.width = 7)
```

This appendix shows how to generate the input files for the climate reconstructions, 
based on the CRU TS4.04 dataset (Harris et al., 2020).

The target output files are:
```{bash}
"cru_ts4.04.1901.2019.cld.dat-clim-1961-1990-int.nc"
"cru_ts4.04.1901.2019.pre.dat-clim-1961-1990-int.nc"
"cru_ts4.04-clim-1961-1990-daily.tmp.nc"
```

## Raw data
Download the following files directly from the CEDA Archive (https://catalogue.ceda.ac.uk/uuid/89e1e34ec3554dc98594a5732622bce9) 
```{r}
ncfiles_raw <- c("cru_ts4.04.1901.2019.cld.dat.nc",
                 "cru_ts4.04.1901.2019.pre.dat.nc",
                 "cru_ts4.04.1901.2019.tmn.dat.nc",
                 "cru_ts4.04.1901.2019.tmx.dat.nc",
                 "cru_ts4.04.1901.2019.vap.dat.nc")
ncfiles_var <- c("cld", "pre", "tmn", "tmx", "vap")
path <- "/path/to/CRU/4.04/"
```

## Convert precipitation from `[mm/month]` to `[mm/day]`
```{r}
codos:::convert_units.m2d(
  filename = file.path(path, "cru_ts4.04.1901.2019.pre.dat.nc"),
  varid = "pre",
  output_filename =
    file.path(output_path, "cru_ts4.04.1901.2019.pre.dat-new.nc")
)
```
##### Output file
```{bash}
"cru_ts4.04.1901.2019.pre.dat-new.nc"
```

## Create monthly climatologies: 1961-1990
```{r}
ncfiles_clim <- c("cru_ts4.04.1901.2019.cld.dat-clim-1961-1990.nc",
                  "cru_ts4.04.1901.2019.pre.dat-clim-1961-1990.nc",
                  "cru_ts4.04.1901.2019.tmn.dat-clim-1961-1990.nc",
                  "cru_ts4.04.1901.2019.tmx.dat-clim-1961-1990.nc",
                  "cru_ts4.04.1901.2019.vap.dat-clim-1961-1990.nc")

message("###### Create monthly climatologies: 1961-1990 ######")
for (i in seq_along(ncfiles_raw)) {
  message("Processing: ", ncfiles_var[i], "...")
  codos::monthly_clim(
    filename = file.path(path, ncfiles_raw[i]),
    varid = ncfiles_var[i],
    s_year = 1961,
    e_year = 1990,
    output_filename = file.path(output_path, ncfiles_clim[i])
  )
}
```

##### Output files
```{bash}
"cru_ts4.04.1901.2019.cld.dat-clim-1961-1990.nc"
"cru_ts4.04.1901.2019.pre.dat-clim-1961-1990.nc"
"cru_ts4.04.1901.2019.tmn.dat-clim-1961-1990.nc"
"cru_ts4.04.1901.2019.tmx.dat-clim-1961-1990.nc"
"cru_ts4.04.1901.2019.vap.dat-clim-1961-1990.nc"
```

## Interpolate monthly data to daily
```{r}
ncfiles_clim <- c("cru_ts4.04.1901.2019.cld.dat-clim-1961-1990.nc",
                  "cru_ts4.04.1901.2019.pre.dat-clim-1961-1990.nc",
                  "cru_ts4.04.1901.2019.tmn.dat-clim-1961-1990.nc",
                  "cru_ts4.04.1901.2019.tmx.dat-clim-1961-1990.nc",
                  "cru_ts4.04.1901.2019.vap.dat-clim-1961-1990.nc")
ncfiles_var <- c("cld", "pre", "tmn", "tmx", "vap")
ncfiles_clim_int <- c("cru_ts4.04.1901.2019.cld.dat-clim-1961-1990-int.nc",
                      "cru_ts4.04.1901.2019.pre.dat-clim-1961-1990-int.nc",
                      "cru_ts4.04.1901.2019.tmn.dat-clim-1961-1990-int.nc",
                      "cru_ts4.04.1901.2019.tmx.dat-clim-1961-1990-int.nc",
                      "cru_ts4.04.1901.2019.vap.dat-clim-1961-1990-int.nc")

message("###### Interpolate monthly data to daily ######")
for (i in seq_along(ncfiles_raw)) {
  message("Processing: ", ncfiles_var[i], "...")
  codos:::nc_int(
    filename = file.path(output_path, ncfiles_clim[i]),
    varid = ncfiles_var[i],
    s_year = 1961,
    cpus = 20,
    output_filename = file.path(output_path, ncfiles_clim_int[i])
  )
}
```

##### Output files
```{bash}
"cru_ts4.04.1901.2019.cld.dat-clim-1961-1990-int.nc"
"cru_ts4.04.1901.2019.pre.dat-clim-1961-1990-int.nc"
"cru_ts4.04.1901.2019.tmn.dat-clim-1961-1990-int.nc"
"cru_ts4.04.1901.2019.tmx.dat-clim-1961-1990-int.nc"
"cru_ts4.04.1901.2019.vap.dat-clim-1961-1990-int.nc"
```

## Calculate daily temperature
```{r}
tmin <- file.path(path, "cru_ts4.04.1901.2019.tmn.dat-clim-1961-1990-int.nc")
tmax <- file.path(path, "cru_ts4.04.1901.2019.tmx.dat-clim-1961-1990-int.nc")
output_filename <- file.path(path, "cru_ts4.04-clim-1961-1990-daily.tmp.nc")
codos::daily_temp(tmin = list(filename = tmin, id = "tmn"),
                  tmax = list(filename = tmax, id = "tmx"),
                  output_filename = output_filename)
```

##### Output file
```{bash}
"cru_ts4.04-clim-1961-1990-daily.tmp.nc"
```


## References
Harris, I., Osborn, T.J., Jones, P., Lister, D., 2020. Version 4 of the CRU TS monthly high-resolution gridded multivariate climate dataset. Sci Data 7, 109. https://doi.org/10.1038/s41597-020-0453-3
